Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-27 arxiv/proof_gaussian_case.tex
sec B wip

The core idea is to study $G(z)$ using the Schur complement formula.
First, we consider the diagonal entries of $G(z)$ for each block in $\cal I_0$, $\cal I_1$, and $\cal I_2$.
For any $i$ in $\cal I_0$, any $\mu$ in $\cal I_1$, and any $\nu$ in $\cal I_2$, we have that
\begin{align*}
	&G_{i,i}^{-1}(z) = -z - \frac{\lambda_i^2}{n} \sum_{\mu,\nu\in \mathcal I_1} Z^{(1)}_{\mu i}Z^{(1)}_{\nu i}G^{\left( i \right)}_{\mu,\nu}(z) - \frac{1}{n} \sum_{\mu,\nu\in \mathcal I_2} Z^{(2)}_{\mu i}Z^{(2)}_{\nu i} G^{\left( i \right)}_{\mu,\nu}(z) -\frac{2\lambda_i}{n} \sum_{\mu\in \cal I_1,\nu\in \mathcal I_2} Z^{(1)}_{\mu i}Z^{(2)}_{\nu i}G^{\left( i \right)}_{\mu,\nu}(z) \\
	&G_{\mu,\mu}^{-1}(z) =  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}\lambda_i \lambda_j Z^{(1)}_{\mu i}Z^{(1)}_{\mu j} G^{\left(\mu\right)}_{ij}(z) \\
	&G_{\nu,\nu}^{-1}(z) =  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}  Z^{(2)}_{\nu i}Z^{(2)}_{\nu j}  G^{\left(\nu\right)}_{i,j}(z).
\end{align*}
For the first equation, we expand the Schur complement formula $G_{ii}^{-1}(z) = -z - H_i G^{(i)}(z) H_{i}^\top$, where $H_i$ is the $i$-th row of $H$ with the $(i,i)$-th entry removed.
The second and third equation follow by similar calculations.

@@

Finally, we derive the matrix limit $\Gi(z)$.
Inserting the approximate identity \eqref{approx m12 add} into equations \eqref{1self_Gii} and \eqref{2.5self_Gmu}, we get that for  $i \in \cal I_0$, $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
$$G_{ii}(z)=(-z +\lambda_i^2 a_{1}(z) + a_2(z)+\oo(1)^{-1},\quad G_{\mu\mu}=-\frac{\rho_1+\rho_2}{\rho_1}a_{1}(z)+\oo(1),\quad G_{\nu\nu}=-\frac{\rho_1+\rho_2}{\rho_2}a_{2}(z)+\oo(1),$$
with overwhelming probability. These explain the diagonal entries of $\Gi$ in equation \eqref{defn_piw}. For the off-diagonal entries, they are close to zero due to concentration. For example, for $i\ne j\in \cal I_1$, by Schur complement formula in equation (\ref{resolvent3}), we have
$$G_{ij}=-G_{ii}\Big({\lambda_i}{n^{-1/2}}\sum_{\mu \in \cal I_1} Z^{(1)}_{\mu i} G^{(i)}_{\mu j} + {n^{-1/2}}\sum_{\mu \in \cal I_2} Z^{(2)}_{\mu i} G^{(i)}_{\mu j} \Big).$$
Using Lemma \ref{largedeviation}, we can show that $n^{-1/2}\sum_{\mu \in \cal I_1} Z^{(1)}_{\mu i} G^{(i)}_{\mu j}$ and $n^{-1/2}\sum_{\mu \in \cal I_2} Z^{(2)}_{\mu i} G^{(i)}_{\mu j}$ are both close to zero. The other off-diagonal entries can be bounded in the same way. The bound on the off-diagonal entries will be proved rigorously in Lemma \ref{Z_lemma}.

\paragraph{Notations.}
We say an event $\Xi$ holds with overwhelming probability if for any constant $D>0$, $\mathbb P(\Xi)\ge 1- n^{-D}$ for large enough $n$. Moreover, we say $\Xi$ holds with overwhelming probability in an event $\Omega$ if for any constant $D>0$, $\mathbb P(\Omega\setminus \Xi)\le n^{-D}$ for large enough $n$.
We frequently use the following notion of stochastic domination, which was first introduced in \cite{Average_fluc} and subsequently used in many works on random matrix theory.
%It greatly simplifies the presentation of the results and their proofs by systematizing statements of the form ``$\xi$ is bounded by $\zeta$ with overwhelming probability up to a small power of $n$".

\begin{definition}[Stochastic domination]\label{stoch_domination}
%\begin{itemize}
%\item[(i)]
Let $\xi\equiv \xi^{(n)}$ and $\zeta\equiv \zeta^{(n)}$ be two $n$-dependent random variables.
\HZ{what is $n$-dependent rv?}
We say $\xi$ is stochastically dominated by $\zeta$, denoted by $\xi\prec \zeta$ or $\xi=\OO_\prec(\zeta)$, if for any (small) constant $\e>0$ and (large) constant $D>0$
\[ \bbP\left(|\xi| >n^\e |\zeta|\right)\le n^{-D}, \]
for large enough $n\ge n_0(\e, D)$. Moreover, if $\xi$ and $\zeta$ depend on a parameter $u$, then we say $\xi$ is stochastically dominated by $\zeta$ uniformly in $u\in \cal U$,  if for any constants $\e,D>0$,
\[\sup_{u\in \cal U}\bbP\left(|\xi(u)|>n^\e |\zeta(u)|\right)\le n^{-D}, \]
for large enough $n\ge n_0(\e, D)$.
%\item[(ii)]
%\end{itemize}
%(i) Let
%\[\xi=\left(\xi^{(n)}(u):n\in\bbN, u\in U^{(n)}\right),\hskip 10pt \zeta=\left(\zeta^{(n)}(u):n\in\bbN, u\in U^{(n)}\right)\]
%be two families of nonnegative random variables, where $U^{(n)}$ is a possibly $n$-dependent parameter set. We say $\xi$ is stochastically dominated by $\zeta$, uniformly in $u$, if for any fixed (small) $\epsilon>0$ and (large) $D>0$,
%\[\sup_{u\in U^{(n)}}\bbP\left[\xi^{(n)}(u)>n^\epsilon\zeta^{(n)}(u)\right]\le n^{-D}\]
%for large enough $n\ge n_0(\epsilon, D)$, and we shall use the notation $\xi\prec\zeta$.
%%Throughout this paper, the stochastic domination will always be uniform in all parameters that are not explicitly fixed (such as matrix indices, and $z$ that takes values in some compact set).
%%Note that $N_0(\epsilon, D)$ may depend on quantities that are explicitly constant, such as $\tau$ in Assumption \ref{assm_big1} and equation \eqref{assm_gap}.
%If for some complex family $\xi$ we have $|\xi|\prec\zeta$, then we will also write $\xi \prec \zeta$ or $\xi=\OO_\prec(\zeta)$.
%\item[(ii)]
%(ii) We extend the definition of $\OO_\prec(\cdot)$ to matrices in the weak operator sense as follows. Let $A$ be a family of random matrices and $\zeta$ be a family of nonnegative random variables. Then $A=O_\prec(\zeta)$ means that $\left|\left\langle\mathbf v, A\mathbf w\right\rangle\right|\prec\zeta \| \mathbf v\|_2 \|\mathbf w\|_2 $ uniformly in any deterministic vectors $\mathbf v$ and $\mathbf w$. Here and throughout the following, whenever we say ``uniformly in any deterministic vectors", we mean that ``uniformly in any deterministic vectors belonging to a set of cardinality $N^{\OO(1)}$".
%\item[(iv)]
%\end{itemize}
\end{definition}

For example, we say that $\xi$ is stochastically dominated by $\zeta$ with overwhelming probability if $\zeta$ is greater than $\xi$ up to a small power of $n^{\e}$.
Since we allow an $n^\e$ factor in the definition of stochastic domination, we have the seemingly strange inequality $(\log n)^C\prec 1$ for any constant $C>0$.
As another example, given a random variable $\xi$, if its moments exist up to any order, then we have $|\xi|\prec 1$. In fact, for any constants $\e, D>0$, by Markov's inequality we have
$$ \P(|\xi|\ge n^{\e})\le n^{-k\e}{\E |\xi|^k}\le n^{-D},$$
if we choose $k$ large enough such that $k\e>D$. As a special case, we have $|\xi|\prec 1$ for a Gaussian random variable $\xi$ with unit variance.

%\HZ{Why is this below part of the def. of stochastic domination?}

>>>>

The core idea is to study $G(z)$ using the Schur complement formula.
First, we consider the diagonal entries of $G(z)$ for each block in $\cal I_0$, $\cal I_1$, and $\cal I_2$.
For any $i$ in $\cal I_0$, any $\mu$ in $\cal I_1$, and any $\nu$ in $\cal I_2$, we have that
\begin{align*}
	&G_{i,i}^{-1}(z) = -z - \frac{\lambda_i^2}{n} \sum_{\mu,\nu\in \mathcal I_1} Z^{(1)}_{\mu i}Z^{(1)}_{\nu i}G^{\left( i \right)}_{\mu,\nu}(z) - \frac{1}{n} \sum_{\mu,\nu\in \mathcal I_2} Z^{(2)}_{\mu i}Z^{(2)}_{\nu i} G^{\left( i \right)}_{\mu,\nu}(z) -\frac{2\lambda_i}{n} \sum_{\mu\in \cal I_1,\nu\in \mathcal I_2} Z^{(1)}_{\mu i}Z^{(2)}_{\nu i}G^{\left( i \right)}_{\mu,\nu}(z) \\
	&G_{\mu,\mu}^{-1}(z) =  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}\lambda_i \lambda_j Z^{(1)}_{\mu i}Z^{(1)}_{\mu j} G^{\left(\mu\right)}_{i,j}(z) \\
	&G_{\nu,\nu}^{-1}(z) =  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}  Z^{(2)}_{\nu i}Z^{(2)}_{\nu j}  G^{\left(\nu\right)}_{i,j}(z).
\end{align*}
For the first equation, we expand the Schur complement formula $G_{ii}^{-1}(z) = -z - H_i G^{(i)}(z) H_{i}^\top$, where $H_i$ is the $i$-th row of $H$ with the $(i,i)$-th entry removed.
The second and third equation follow by similar calculations.

@@

\paragraph{Matrix limit.}
Finally, we derive the matrix limit $\Gi(z)$.
Applying our solution for $m_1(z)$ and $m_2(z)$ into equation \eqref{1self_Gii}, we get that for $i$ in $\cal I_0$,
$G_{i, i}(z) = (-z +\lambda_i^2 a_{1}(z) + a_2(z)+\oo(1))^{-1}$.
Similarly, for $\mu$ in $\cal I_1$, we have that $G_{\mu, \mu}(z) = -\frac{n_1 + n_2}{n_1}a_{1}(z)+\oo(1)$, and for $\nu$ in $\cal I_2$, we have that $G_{\nu, \nu}(z) = -\frac{n_1 + n_2}{n_2}a_{2}(z)+\oo(1)$.
Hence we have derived the diagonal entries of $\Gi(z)$.
In Lemma \ref{Z_lemma}, we show that the off-diagonal entries are close to zero.
For example, for $i\ne j\in \cal I_1$, by Schur complement, we have that
$$G_{i,j}(z) = -G_{i,i}(z)\cdot {n^{-1/2}}\Big({\lambda_i}\sum_{\mu \in \cal I_1} Z^{(1)}_{\mu i} G^{(i)}_{\mu, j}(z) + \sum_{\mu \in \cal I_2} Z^{(2)}_{\mu, i} G^{(i)}_{\mu, j}(z) \Big).$$
Using standard concentration inequalities, we show that $\sum_{\mu \in \cal I_1} Z^{(1)}_{\mu i} G^{(i)}_{\mu, j}(z)$ and $\sum_{\mu \in \cal I_2} Z^{(2)}_{\mu i} G^{(i)}_{\mu, j}(z)$ are both close to zero.
The other off-diagonal entries are bounded similarly.
%The bound on the off-diagonal entries will be proved rigorously .

\paragraph{Notations.}
We introduce several notations for the proof of Theorem \ref{thm_main_RMT}.
We say that an event $\Xi$ holds with overwhelming probability if for any constant $D>0$, $\mathbb P(\Xi)\ge 1- n^{-D}$ for large enough $n$.
Moreover, we say $\Xi$ holds with overwhelming probability in an event $\Omega$ if for any constant $D>0$, $\mathbb P(\Omega\setminus \Xi)\le n^{-D}$ for large enough $n$.
The following notion of stochastic domination, which was first introduced in \cite{Average_fluc}, is commonly used in the study of random matrices.

\begin{definition}[Stochastic domination]\label{stoch_domination}
Let $\xi\equiv \xi^{(n)}$ and $\zeta\equiv \zeta^{(n)}$ be two $n$-dependent random variables.
\HZ{what is $n$-dependent rv?}
We say that $\xi$ is stochastically dominated by $\zeta$, denoted by $\xi\prec \zeta$ or $\xi=\OO_\prec(\zeta)$, if for any small constant $\e > 0$ and any large constant $D > 0$, there exists a function $n_0(\e, D)$ such that for any $n > n_0(\e, D)$,
\[ \bbP\left(|\xi| >n^\e |\zeta|\right)\le n^{-D}. \]
In case $\xi(u)$ and $\zeta(u)$ is a function of $u$ supported in $\cal U$, then we say $\xi(u)$ is stochastically dominated by $\zeta(u)$ uniformly in $\cal U$ if % for any constants $\e, D>0$,
\[ \sup_{u\in \cal U}\bbP\left(|\xi(u)|>n^\e |\zeta(u)|\right)\le n^{-D}. \]
\end{definition}

%For example, we say that $\xi$ is stochastically dominated by $\zeta$ with overwhelming probability if $\zeta$ is greater than $\xi$ up to a small power of $n^{\e}$.
We make several remarks.
First, since we allow an $n^\e$ factor in stochastic domination, we can ignore log factors without loss of generality since $(\log n)^C\prec 1$ for any constant $C>0$.
Second, given a random variable $\xi$ whose moments exist up to any order, we have that $|\xi|\prec 1$.
This is because by Markov's inequality, let $k$ be greater than $D/ \e$, then we have that
$$ \P(|\xi|\ge n^{\e})\le n^{-k\e}{\E |\xi|^k}\le n^{-D}.$$
This implies that a Gaussian random variable $\xi$ with unit variance satisfies that $|\xi|\prec 1$.
